The binary liquid model for gravity-driven microchannel simulations

### 2011-01-04 Text/paper.tex
Finalized my revisions. Sasha sent the figure.
Started doing modifications, but found that didn't merge previous ones.

\begin{abstract}
This work shows that the free-energy binary liquid lattice Boltzmann method is able to
reproduce the gravity driven microchannel flow of an air bubble through liquid. In the low capillary
number
$\Ca$ regime (defined in Eq. \ref{capillary:number:definition}), the flow is thoroughly described by
Bretherton \cite{bretherton},
with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
from the Bretherton correlations for $Ca\geq 0.005$, studied experimentally and numerically. This
work aims to examine the lattice Boltzmann binary liquid model feasibility to obtain valid results
for an air bubble flow in liquid for a range of $Ca$ number. The binary
liquid model is a continuous-interface multiphase model, which raises
the question of how to properly resolve the interface in comparison with the film thickness.
It was found that the film interface needs to be as $50$
percent in comparison with the film thickness for simulation to give convergent results. While
the convergency criteria is fulfilled the results are of good comparison with the
numerical simulations of \citet{giavedoni-numerical} in the range of $Ca$ from $0.01$ to $1$. 
\end{abstract}

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow describes the flow of long bubbles in
microchannels. It was found that the front meniscus film width is proportional
to $\Ca^{2/3}$ for capillary numbers less than $0.003$, where the capillary number is defined as:
\begin{equation}
\label{capillary:number:definition}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is a propagation bubble velocity, and
$\gamma$ is the surface tension between air and liquid.
While there are number
of works which simulate the Bretherton problem with the boundary value method
\cite{ingham-plates,heil-bretherton}, those methods are of limited applicability for problems
involving complex geometries, free interface motion or for problems involving coalescence and/or
droplet breakup. The continuous interface models are more flexible for those kind of simulations.
However, if the interface is spread over several grid nodes, the question of
proper film width resolution in comparison with the interface arises.  This work
concentrates on resolving the parameter range of the binary liquid lattice
Boltzmann method to validate it in the range of capillary numbers.

@@

This work examines the binary liquid free-energy LB model \cite{swift}. The
model simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for air and liquid, which are of largely
different densities and viscosities, it was indicated \cite{bretherton} that inertia effects were
neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
negligible Reynolds number effect on the film thickness for the range of Reynolds numbers.
Therefore, the governing factor for microchannel flows is not the densities ratio, but the
viscosities ratio. Moreover, the results based on the uniform density, as in the case of the LBM
binary liquid model, are of good comparison with other simulations
\cite{giavedoni-numerical,heil-bretherton} taking into the account inertia effects.

@@

\section{Bretherton bubble flow}
Bretherton \cite{bretherton} studied a long bubble
moving in a tube filled with liquid. It was found that the deposition film width
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later on
it was realized \cite{wong-films,wong-pressure} that the film deposition
is proportional to $\Ca^{2/3}$ only in the specific region behind the front meniscus and
that the film width
varies over the bubble length for bubbles of finite length. Numerical
simulations \cite{giavedoni-numerical} and experimental studies
\cite{kreutzer-pressure-drop} showed as well a Reynolds number dependence, and
a deviation from the $\Ca^{2/3}$ rule for capillary numbers larger than $0.003$.
To be able to predict consistently the flow pattern for capillaries in
different ranges of parameters, simulations have been developed and are typically validated with the
small capillary numbers Bretherton problem.

There are a number of numerical methods which
were used for the simulation of the Taylor/Bretherton flow.
\citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillary using the finite volume method.
\citet{kreutzer-pressure-drop} used the finite volume method as well to perform
simulations of a circular capillary for a number of different
Reynolds and capillary numbers. \citeauthor{wong-films} \cite{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and showed the derivations of different shapes in the
slug crossections and menisci appearance.
\citet{heil-bretherton,ingham-plates} studied air finger propagation in
a two-dimensional channel for a range of Reynolds and capillary
numbers using finite element method. \citet{giavedoni-numerical} performed crossvalidation of the
finite element solution developed by them with previously published results.
The solution is available for circular and planar cases.

@@

\section{Lattice Boltzmann benchmark}
To properly simulate and to compare simulation results with the data
published in the literature, one needs to design a numerical benchmark addressing
the following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive nature of the interface in the multiphase model should not
influence the film width results. This is obviously achieved only for certain
resolutions of the interface in comparison with the film width. The same applies to spurious
currents arising from surface tension discretization which
can influence mass transfer in the thin film, thereby compromising the
overall solution. Therefore the diffusive interface should be of nengligible influence on the film
thickness.
 \item The wettability coefficient which in principle can control the
dynamic angle and the menisci appearance \cite{pagonabarraga-finger} should not
influence the overall solution.
 \item The free-energy model examined in the work is a single density, two
viscosities model. The classical Bretherton problem is a free-interface problem between liquid and
air
where the film thickness is established by itself. Thus, the viscosity and density ratios for the
classical
Bretherton problem should
be large. However, in terms of the lattice Boltzmann binary liquid framework only the viscosities
ratio can be addressed. The results show that the results are consistent with other works and
inertia effects can be neglected.
  \item The Bretherton problem addresses a bubble flow driven by a pressure
difference, not a body force. However, to the best authors knowledge the pressure boundaries
conditions are still not yet developed for the binary liquid model, because it involves the
treatment of associated gradients and laplacians on the boundary.
This kind of pressure boundary conditions for the binary liquid model is currently
under development, and preliminary results show that the differences between
pressure difference flows and flows driven by a body force are small as far as the film
deposition thickness is of interest.  In this work we limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability. The case of
body force driven flows implies the periodic boundary conditions. As soon as the periodic boundary
conditions are applied, not one bubble but whole bubble train is simulated. In this case one needs
to insure that bubbles do not influence each other making large distance between them.
 \item Most simulations are done for a channel with circular crosssection,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate the Laplacian and gradients near boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
the analysis of error introduced by the boundary conditions becomes an issue. Though the circular
axis-symmetric case can be simulated on rectangular grid by introducing specific mass and force
terms in the continuity and Navier-Stokes equation \cite{halliday-circular}, the lattice Boltzmann
incorporation of force and mass terms in the process of development and validation. Therefore
the benchmark is chosen as a plane-symmetric two-dimensional case.

@@

We take the bubble length as $5$ channel widths
  and the distance between bubbles is $2$ bubble lengths. This gives a
grid size of $400 \times 6000$. In terms of computer memory
requirements, the calculations with double precision accuracy demand:
\begin{equation}
\begin{aligned}
&2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
} \times \,\, 9 \text{ number of
populations}\\
&\times\,400\,\times\,6000\approx 345 \mathrm{MB}. 
\end{aligned}
\end{equation}
Even with this naive assumption for the $62.5$ percent resolution of the interface width of the film
width, one can get the memory
requirement of $345 \mathrm{MB}$.

@@

 \item[II $\bm{Ca=0.05}$] 
   The parameters in the previous subsection imply long
simulations for system to come to equilibrium and this number increase with decrease of the
capillary number. Bretherton obtained the analytical solution for the range of low capillary numbers
less than
  $0.005$. However, a lot of simulations extended it to the range of Capillary numbers greater than
$0.005$, and to higher Reynolds numbers. One can refer to figures
  \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from the work of
  \citet{giavedoni-numerical,heil-bretherton}, who crossvalidated their
  simulations with Bretherton and each other. Even though the LBM can simulate
flows with small capillary numbers, i.e. less than $0.005$, the simulation time
is relatively long. Therefore, we focus our work on the simulations of
\citet{giavedoni-numerical} who crossvalidated
results for all ranges of capillary numbers.  Our intention is
  to show proper physics behavior for moderate parameters
which
  can be easily and quickly validated on a computer.
  One can see from Fig. \ref{fig:heil:planar} that the width for the
  Capillary number $0.05$ ranges from $0.12$ to $0.15$, while the Reynolds
  number ranges from $0$ to $300$.

  The film width for such small Capillary numbers is taken from
  \ref{fig:heil:planar}, and for Reynolds number $40$ equals $0.12$. We want to
  resolve $6\%$ of whole channel width with $12$ lattice units, where $5$
lattice units are related to the interface resolution.  Therefore the channel
  width is $200$ lattice units. Therefore using the assumptions for the benchmark 
the grid size is $200 \times 3000$. The velocity can be
calculated from the Capillary number as:
  \begin{equation}
  U_{\mathrm{bubble}}=0.05 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.05 \frac{0.037712}{0.6666}=2.82\cdot
  10^{-3}.
  \end{equation}
  The velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution:
  \begin{equation}
  \begin{aligned}
  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma Ca=3.77\cdot10^{-7}.
  \end{aligned}
  \end{equation}
  Our approach to the initial conditions is  to take values from the previously performed
numerical simulations and to calculate the grid sizes from them, matching the
body force gradient through the solution of Poiseuille flow.

@@

\subsection{Grid refinement}
To properly estimate the interface resolution one needs to study the convergence
of the results as a function of the grid resolution. To do that, all parameters are held fixed,
including the bubble velocity, the capillary number, and only the grid
resolution is changed. Our goal is to determine the ratio of the interface width to 
film thickness where results are not dependent on the grid resolution.

@@

After choosing the reference parameters, the grid refinement procedure
needs to keep the macroscopic parameters constant.  It is easy to check
that it yields the following quantity to
be independent of grid size
$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
x } = \mathrm{const}$. We performed a number of simulations for the prescribed grids. The
simulations in terms of grid numbers $N_x$, $N_y$, the film thickness $\delta$, center bubble
velocity $U_{bubble}$ are summarized in Table \ref{table:parameters:grid:refinement}.
The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
see the visual results converge for $H_{\mathrm{eff}}\geq 175$. However, one can see
that with proper initialization techniques, large enough time and the wall
wettability, results are different only in the 3rd digit even for
underresolved film widths. To calculate how well the interface thickness is
resolved, the ratio of the interface thickness to the film width thickness is calculated. The
interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. The ratio of the interface width to the film thickness $5\xi/H_{\mathrm{film}}$
 is shown in Table \ref{table:parameters:grid:refinement}. Based on results one can conclude that
 the interface needs to be resolved as $40-50$ percent of the
expected film width for simulations to be convergent. We further examine the
velocities in the center of the bubble to calculate the Capillary number. One can see from Table
\ref{table:parameters:grid:refinement} that the bubble velocities are pretty consistent and the
calculated capillary number corresponding to those values is $0.07$, which is slightly
larger than the expected capillary number. The corresponding error can be
attributed to the density and viscosity ratios which are much larger in the
Bretherton problem and Poiseuille profile body force initialization.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Grid/norm_grid_profs.eps}
\caption{Grid-refined profiles for the effective
channel widths
$H_{\mathrm{eff}}=100,125,150,175,200,225$.$\delta$ is scaled on $H_{eff}$ and $\delta=0$
corresponds to the wall location. The profiles were taken at $x=14$ nondimensional coordinate.
Capillary number is $Ca=0.05$. Capillary number obtained from simulations $Ca=0.07$. One can
see
convergent results even for the underresolved film
thickness.\label{fig:grid:profiles}}
\end{figure}
%Here will be a consolidated table
\begin{table}
\begin{tabularx}{0.97\textwidth}{|X|X|X|X|X|X|}
\hline
$N_x$&$N_y$&$\delta$&$U_{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_{iter}$\\
\hline
$1500$&$102$&$0.0694$&$0.0041$&$0.824$&$200000$\\
\hline
$1875$&$127$&$0.0688$&$0.0041$&$0.646$&$250000$\\
\hline
$2250$&$152$&$0.0676$&$0.0040$&$0.539$&$300000$\\
\hline
$2625$&$177$&$0.0679$&$0.0040$&$0.453$&$350000$\\
\hline
$3000$&$202$&$0.0668$&$0.0041$&$0.400$&$400000$\\
\hline
$3375$&$227$&$0.0663$&$0.0039$&$0.355$&$450000$\\
\hline
\end{tabularx}
\caption{The parameters and results for studying grid resolution simulations. Domain simulated is
of size $N_x\, \mathrm{x}\, N_y$. $U_{bubble}$ is center
bubble velocity measured at the scaled to $H_eff$ $x=14.0$.
$5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of the film in
lattice Boltzmann units. $Ca=0.05$ was taken to initialize the simulation.
\label{table:parameters:grid:refinement}}
\end{table}

\subsection{The influence of the wall gradient}
The simulation results should not depend on the wall wettability. Wettability is defined
through the phase gradient \cite{pooley-contact}, $\partial_n \phi$.  We took a large
enough grid resolution in order for simulations to be consistent -- the grid size was
$177 \times 2626$ and the initial film width was
$12$ lattice boltzmann units, which corresponds to the predicted film thickness.
We examined $11$ different values for the wall
gradient ranging from $-1$ to $1$. The parameters are summarized in Table
\ref{table:parameters:wall:gradient}. The results are consistent among all the
wall gradients given that the interface is resolved in the right way. For 
given wall gradients the values of interface thickness values are of $0.2\%$ relative accuracy.  For
the wall gradients $0.8$
and $1.0$ the
simulations are unstable. One can see the simulated profiles in Fig.
\ref{fig:gradients:profiles}. The latter shows that the negative wall gradient values cause 
unphysical phase values more than $1.0$ near the wall. One should attribute them to the numerical
imposement of wall gradients. Therefore, they do not spoil numerical scheme as whole. Note that
negative values of the phase gradient are
preferable, since the phase of the liquid adjacent to the wall has the value of $1$.
The phase values near the wall are above $1$ and do not interact
with the bubble which has phase value $-1$.  In case of positive
gradients the values near the wall are below $1$, and the gradient profile can
fuse with the values of the gas phase. This case is attributed when to the slug flow when gas
touches the wall.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Wall/phase_grad_profiles.eps}
\caption{Phase profiles for different wall gradients. One can see that if the width
is properly resolved then the wettability of the wall does affect the film thickness. The
simulations were conducted for $200,000$ iterations. The measured profile was taken at the
nondimensional coordinate scaled to $H_{\mathrm{eff}}$ as $x=11.42$. $Ca=0.05$ was taken to
initialize simulation.
\label{fig:gradients:profiles}
}
\end{figure}

@@

\subsection{The influence of different viscosities}
The results obtained earlier are taken for the liquid viscosity ratio $10$.
We performed the same simulations for the liquid viscosity ratio of $20$.
The relaxation parameters were taken as $\tau_{\mathrm{liquid}}=4.5$ and $\tau_{\mathrm{bubble}}=0.7$. We
did not find any significant differences with the results obtained earlier. The
comparison for the liquid thicknesses is presented in Fig.
\ref{fig:capillary:viscous}. The results validate our assumption that in
the low capillary flow regime the density ratio does not affect the film thickness,
and that the gas-liquid viscosity ratio of the order $10$ is enough to obtain results
consistent with the Bretherton/Taylor problem.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Capillary_Viscous/capillaries_viscous.eps}
\caption{The film widths versus capillary number for
$\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=10$ and $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=20$. The
results show good consistency and mutual agreement.\label{fig:capillary:viscous}}
\end{figure}

>>>>

\begin{abstract}
This work shows that the free-energy binary liquid lattice Boltzmann method is able to
reproduce the gravity driven microchannel flow of a gas bubble through liquid. In the regime of low
capillary
numbers
$\Ca$ (defined in Eq. \ref{capillary:number:definition}), the flow is thoroughly described by
Bretherton \cite{bretherton},
with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
from the Bretherton correlation for $Ca\geq 0.005$, studied experimentally and numerically. This
work aims to examine the lattice Boltzmann binary liquid model feasibility to obtain valid results
for a bubble flow in liquid for a range of $Ca$ number. The binary
liquid model is a continuous-interface multiphase model allowing to properly resolve the interface.
It was found that the film interface needs to be as $50$
percent in comparison with the film thickness for simulation to give convergent results. While
the convergency criteria is fulfilled the results are of good comparison with the
numerical simulations of \citet{giavedoni-numerical} in the range of $Ca$ from $0.01$ to $1$. 
\end{abstract}

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow describes the flow of long bubbles in
microchannels. It was found that the front meniscus film thickness is proportional
to $\Ca^{2/3}$ for capillary numbers less than $0.003$, where the capillary number is defined as:
\begin{equation}
\label{capillary:number:definition}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is a propagation bubble velocity, and
$\gamma$ is the surface tension between gas and liquid.
While there are number
of works which simulate the Bretherton problem with the boundary value method
\cite{ingham-plates,heil-bretherton}, those methods are of limited applicability for problems
involving complex geometries, free interface motion or for problems involving coalescence and/or
droplet breakup. The continuous interface models are more flexible for those kind of simulations.
However, if the interface is spread over several grid nodes, the question of
proper film thickness resolution in comparison with the interface arises.  This work
concentrates on resolving the parameter range of the binary liquid lattice
Boltzmann method to validate it in the range of capillary numbers.

@@

This work examines the binary liquid free-energy LB model \cite{swift}. The
model simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for gas and liquid, which are of largely
different densities and viscosities, it was indicated \cite{bretherton} that inertia effects were
neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
negligible Reynolds number effect on the film thickness for the range of Reynolds numbers.
Therefore, the governing factor for microchannel flows is not the densities ratio, but the
viscosities ratio. Moreover, the results based on the uniform density, as in the case of the LBM
binary liquid model, are of good comparison with other simulations
\cite{giavedoni-numerical,heil-bretherton} taking into the account inertia effects.

@@

\section{Bretherton bubble flow}
Bretherton \cite{bretherton} studied a long bubble
moving in a tube filled with liquid. It was found that the deposition film thickness
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later on
it was realized \cite{wong-films,wong-pressure} that the film deposition
is proportional to $\Ca^{2/3}$ only in the specific region behind the front meniscus and
that the film thickness
varies over the bubble length for bubbles of finite length. Numerical
simulations \cite{giavedoni-numerical} and experimental studies
\cite{kreutzer-pressure-drop} showed as well a Reynolds number dependence, and
a deviation from the $\Ca^{2/3}$ rule for capillary numbers larger than $0.003$.
To be able to predict consistently the flow pattern for capillaries in
different ranges of parameters, simulations have been developed and are typically validated with the
small capillary numbers Bretherton problem.

There are a number of numerical methods which
were used for the simulation of the Taylor/Bretherton flow.
\citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillary using the finite volume method.
\citet{kreutzer-pressure-drop} used the finite volume method as well to perform
simulations of a circular capillary for a number of different
Reynolds and capillary numbers. \citeauthor{wong-films} \cite{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and showed the derivations of different shapes in the
slug crossections and menisci appearance.
\citet{heil-bretherton,ingham-plates} studied gas finger propagation in
a two-dimensional channel for a range of Reynolds and capillary
numbers using finite element method. \citet{giavedoni-numerical} performed crossvalidation of the
finite element solution developed by them with previously published results.
The solution is available for circular and planar cases.

@@

\section{Lattice Boltzmann benchmark}
To properly simulate and to compare simulation results with the data
published in the literature, one needs to design a numerical benchmark addressing
the following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive nature of the interface in the multiphase model should not
influence the film thickness results. This is obviously achieved only for certain
resolutions of the interface in comparison with the film thickness. The same applies to spurious
currents arising from surface tension discretization which
can influence mass transfer in the thin film, thereby compromising the
overall solution. Therefore the diffusive interface should be of nengligible influence on the film
thickness.
 \item The wettability coefficient which in principle can control the
dynamic angle and the menisci appearance \cite{pagonabarraga-finger} should not
influence the overall solution.
 \item The free-energy model examined in the work is a single density, two
viscosities model. The classical Bretherton problem is a free-interface problem between liquid and
gas
where the film thickness is established by itself. Thus, the viscosity and density ratios for the
classical
Bretherton problem should
be large. However, in terms of the lattice Boltzmann binary liquid framework only the viscosities
ratio can be addressed. The results show that the results are consistent with other works and
inertia effects can be neglected.
  \item The Bretherton problem addresses a bubble flow driven by a pressure
difference, not a body force. However, to the best authors knowledge the pressure boundaries
conditions are still not yet developed for the binary liquid model, because it involves the
treatment of associated gradients and laplacians on the boundary.
This kind of pressure boundary conditions for the binary liquid model is currently
under development, and preliminary results show that the differences between
pressure difference flows and flows driven by a body force are small as far as the film
deposition thickness is of interest.  In this work we limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability. The case of
body force driven flows implies the periodic boundary conditions. As soon as the periodic boundary
conditions are applied, not one bubble but whole bubble train is simulated. In this case one needs
to insure that bubbles do not influence each other making large distance between them.
 \item Most simulations are done for a channel with circular crosssection,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate the Laplacian and gradients near boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
the analysis of error introduced by the boundary conditions becomes an issue. Though the circular
axis-symmetric case can be simulated on rectangular grid by introducing specific mass and force
terms in the continuity and Navier-Stokes equation \cite{halliday-circular}, the lattice Boltzmann
incorporation of force and mass terms in the process of development and validation. Therefore
the benchmark is chosen as a plane-symmetric two-dimensional case.

@@

We take the bubble length as $5$ channel widths
  and the distance between bubbles is $2$ bubble lengths. This gives a
grid size of $400 \times 6000$. In terms of computer memory
requirements, the calculations with double precision accuracy demand:
\begin{equation}
\begin{aligned}
&2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
} \times \,\, 9 \text{ number of
populations}\\
&\times\,400\,\times\,6000\approx 345 \mathrm{MB}. 
\end{aligned}
\end{equation}
Even with this naive assumption for the $62.5$ percent resolution of the interface thickness of the
film thickness, one can get the memory
requirement of $345 \mathrm{MB}$.

@@

 \item[II $\bm{Ca=0.05}$] 
   The parameters in the previous subsection imply long
simulations for system to come to equilibrium and this number increase with decrease of the
capillary number. Bretherton obtained the analytical solution for the range of low capillary numbers
less than
  $0.005$. However, a lot of simulations extended it to the range of Capillary numbers greater than
$0.005$, and to higher Reynolds numbers. One can refer to figures
  \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from the work of
  \citet{giavedoni-numerical,heil-bretherton}, who crossvalidated their
  simulations with Bretherton and each other. Even though the LBM can simulate
flows with small capillary numbers, i.e. less than $0.005$, the simulation time
is relatively long. Therefore, we focus our work on the simulations of
\citet{giavedoni-numerical} who crossvalidated
results for all ranges of capillary numbers.  Our intention is
  to show proper physics behavior for moderate parameters
which
  can be easily and quickly validated on a computer.
  One can see from Fig. \ref{fig:heil:planar} that the film thickness for the
  Capillary number $0.05$ ranges from $0.12$ to $0.15$, while the Reynolds
  number ranges from $0$ to $300$.

  The film thickness for such small Capillary numbers is taken from
  \ref{fig:heil:planar}, and for Reynolds number $40$ equals $0.12$. We want to
  resolve $6\%$ of whole channel width with $12$ lattice units, where $5$
lattice units are related to the interface resolution.  Therefore the channel
  width is $200$ lattice units. Therefore using the assumptions for the benchmark 
the grid size is $200 \times 3000$. The velocity can be
calculated from the Capillary number as:
  \begin{equation}
  U_{\mathrm{bubble}}=0.05 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.05 \frac{0.037712}{0.6666}=2.82\cdot
  10^{-3}.
  \end{equation}
  The velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution:
  \begin{equation}
  \begin{aligned}
  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma Ca=3.77\cdot10^{-7}.
  \end{aligned}
  \end{equation}
  Our approach to the initial conditions is  to take values from the previously performed
numerical simulations and to calculate the grid sizes from them, matching the
body force gradient through the solution of Poiseuille flow.

@@

\subsection{Grid refinement}
To properly estimate the interface resolution one needs to study the convergence
of the results as a function of the grid resolution. To do that, all parameters are held fixed,
including the bubble velocity, the capillary number, and only the grid
resolution is changed. Our goal is to determine the ratio of the interface thickness to 
film thickness where results are not dependent on the grid resolution.

@@

After choosing the reference parameters, the grid refinement procedure
needs to keep the macroscopic parameters constant.  It is easy to check
that it yields the following quantity to
be independent of grid size
$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
x } = \mathrm{const}$. We performed a number of simulations for the prescribed grids. The
simulations in terms of grid numbers $N_x$, $N_y$, the film thickness $\delta$, center bubble
velocity $U_{bubble}$ are summarized in Table \ref{table:parameters:grid:refinement}.
The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
see the visual results converge for $H_{\mathrm{eff}}\geq 175$. However, one can see
that with proper initialization techniques, large enough time and the wall
wettability, results are different only in the 3rd digit even for
underresolved film thicknesses. To calculate how well the interface thickness is
resolved, the ratio of the interface thickness to the film thickness is calculated. The
interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. The ratio of the interface thickness to the film thickness
$5\xi/H_{\mathrm{film}}$
 is shown in Table \ref{table:parameters:grid:refinement}. Based on results one can conclude that
 the interface needs to be resolved as $40-50$ percent of the
expected film thickness for simulations to be convergent. We further examine the
velocities in the center of the bubble to calculate the Capillary number. One can see from Table
\ref{table:parameters:grid:refinement} that the bubble velocities are pretty consistent and the
calculated capillary number corresponding to those values is $0.07$, which is slightly
larger than the expected capillary number. The corresponding error can be
attributed to the density and viscosity ratios which are much larger in the
Bretherton problem and Poiseuille profile body force initialization.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Grid/norm_grid_profs.eps}
\caption{Grid-refined profiles for the effective
channel widths
$H_{\mathrm{eff}}=100,125,150,175,200,225$.$\delta$ is scaled on $H_{eff}$ and $\delta=0$
corresponds to the wall location. The profiles were taken at $x=14$ nondimensional coordinate.
Capillary number is $Ca=0.05$. Capillary number obtained from simulations $Ca=0.07$. One can
see
convergent results even for the underresolved film
thickness.\label{fig:grid:profiles}}
\end{figure}
%Here will be a consolidated table
\begin{table}
\begin{tabularx}{0.97\textwidth}{|X|X|X|X|X|X|}
\hline
$N_x$&$N_y$&$\delta$&$U_{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_{iter}$\\
\hline
$1500$&$102$&$0.0694$&$0.0041$&$0.824$&$200000$\\
\hline
$1875$&$127$&$0.0688$&$0.0041$&$0.646$&$250000$\\
\hline
$2250$&$152$&$0.0676$&$0.0040$&$0.539$&$300000$\\
\hline
$2625$&$177$&$0.0679$&$0.0040$&$0.453$&$350000$\\
\hline
$3000$&$202$&$0.0668$&$0.0041$&$0.400$&$400000$\\
\hline
$3375$&$227$&$0.0663$&$0.0039$&$0.355$&$450000$\\
\hline
\end{tabularx}
\caption{The parameters and results for studying grid resolution simulations. Domain simulated is
of size $N_x\, \mathrm{x}\, N_y$. $U_{bubble}$ is center
bubble velocity measured at the scaled to $H_eff$ $x=14.0$.
$5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of the film in
lattice Boltzmann units. $Ca=0.05$ was taken to initialize the simulation.
\label{table:parameters:grid:refinement}}
\end{table}

\subsection{The influence of the wall gradient}
The simulation results should not depend on the wall wettability. Wettability is defined
through the phase gradient \cite{pooley-contact}, $\partial_n \phi$.  We took a large
enough grid resolution in order for simulations to be consistent -- the grid size was
$177 \times 2626$ and the initial film thickness was
$12$ lattice boltzmann units, which corresponds to the predicted film thickness.
We examined $11$ different values for the wall
gradient ranging from $-1$ to $1$. The parameters are summarized in Table
\ref{table:parameters:wall:gradient}. The results are consistent among all the
wall gradients given that the interface is resolved in the right way. For 
given wall gradients the values of interface thickness values are of $0.2\%$ relative accuracy.  For
the wall gradients $0.8$
and $1.0$ the
simulations are unstable. One can see the simulated profiles in Fig.
\ref{fig:gradients:profiles}. The latter shows that the negative wall gradient values cause 
unphysical phase values more than $1.0$ near the wall. One should attribute them to the numerical
imposement of wall gradients. Therefore, they do not spoil numerical scheme as whole. Note that
negative values of the phase gradient are
preferable, since the phase of the liquid adjacent to the wall has the value of $1$.
The phase values near the wall are above $1$ and do not interact
with the bubble which has phase value $-1$.  In case of positive
gradients the values near the wall are below $1$, and the gradient profile can
fuse with the values of the gas phase. This case is attributed when to the slug flow when gas
touches the wall.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Wall/phase_grad_profiles.eps}
\caption{Phase profiles for different wall gradients. One can see that if the thickness
is properly resolved then the wettability of the wall does affect the film thickness. The
simulations were conducted for $200,000$ iterations. The measured profile was taken at the
nondimensional coordinate scaled to $H_{\mathrm{eff}}$ as $x=11.42$. $Ca=0.05$ was taken to
initialize simulation.
\label{fig:gradients:profiles}
}
\end{figure}

@@

\subsection{The influence of different viscosities}
The results obtained earlier are taken for the liquid viscosity ratio $10$.
We performed the same simulations for the liquid viscosity ratio of $20$.
The relaxation parameters were taken as $\tau_{\mathrm{liquid}}=4.5$ and $\tau_{\mathrm{bubble}}=0.7$. We
did not find any significant differences with the results obtained earlier. The
comparison for the liquid thicknesses is presented in Fig.
\ref{fig:capillary:viscous}. The results validate our assumption that in
the low capillary flow regime the density ratio does not affect the film thickness,
and that the gas-liquid viscosity ratio of the order $10$ is enough to obtain results
consistent with the Bretherton/Taylor problem.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Capillary_Viscous/capillaries_viscous.eps}
\caption{The film thicknesses versus capillary number for
$\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=10$ and $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=20$. The
results show good consistency and mutual agreement.\label{fig:capillary:viscous}}
\end{figure}
